---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
df <- read.csv("activity.csv", header = TRUE)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
sum_steps <- tapply(df$steps, df$date, sum, simplify = TRUE, na.rm = TRUE)
barplot(sum_steps, main = "The total number of steps taken each day")
```

The mean of the total number of steps taken per day
```{r, echo=TRUE}
mean(sum_steps, na.rm = TRUE)
```

The median of the total number of steps taken per day
```{r, echo=TRUE}
median(sum_steps, na.rm = TRUE)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
mean_intervals <- tapply(df$steps, df$interval, mean, simplify = TRUE, na.rm = TRUE)
plot(mean_intervals, type = "l", xlab = "5-minute interval", ylab = "the average number of steps")
```

5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r, echo=TRUE}
names(which.max(mean_intervals))
```

## Imputing missing values
The total number of missing values in the dataset
```{r, echo=TRUE}
nrow(subset(df, is.na(df$steps)))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in to __the mean for that 5-minute interval__
```{r, echo=TRUE}
df_filled <- df
for (i in 1:nrow(df_filled)) {
	if (is.na(df_filled[i,1])) {
		df_filled[i,1] <- mean_intervals[as.character(df_filled[i,3])]
	}
}
sum_steps_filled <- tapply(df_filled$steps, df_filled$date, sum, simplify = TRUE, na.rm = TRUE)
barplot(sum_steps_filled, main = "The total number of steps taken each day")
```

The mean of the total number of steps taken per day
```{r, echo=TRUE}
mean(sum_steps_filled, na.rm = TRUE)
```

The median of the total number of steps taken per day
```{r, echo=TRUE}
median(sum_steps_filled, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
weekdays_dir <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
df_filled$day <- factor((weekdays(as.Date(df_filled$date)) %in% weekdays_dir), labels=c('weekend', 'weekday'))
```

```{r, echo=TRUE}
par(mfrow=c(2,1))
weekdays <- subset(df_filled, df_filled$day == "weekday")
plot(tapply(weekdays$steps, weekdays$interval, mean, simplify = TRUE),
	type = "l", xlab = "5-minute interval", ylab = "the average number of steps", main="weekdays")

weekends <- subset(df_filled, df_filled$day == "weekend")
plot(tapply(weekends$steps, weekends$interval, mean, simplify = TRUE),
	type = "l", xlab = "5-minute interval", ylab = "the average number of steps", main="weekends")
```